### 2025-08-06 actuarial.qmd
20250806 - addition and misspelling

As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less finanical resources than its competitors—to compete with teams that hard more money to sign players.
Because the team did not have the resources to sign the best players, they had to find to find other ways to find the optimal players that they could afford.
So, he used statistical formulas that weight variables, such as on-base percentage and slugging percentage, according to their value for winning games, for the use of selecting players.
His approach became well-known based on the Michael Lewis book, "Moneyball: The Art of Winning an Unfair Game", and the eventual movie, "Moneyball".

@@

There is sometimes a misconception that formulas cannot account for qualitative information.
However, that is not true.
Qualitative information can be scored or coded to be quantified so that it can be included in statistical formulas.
For instance, if an expert scout is able to meaningfully assess a player's [cognitive and motivational factors](#sec-evalCogMotivational) (i.e., the "[X factor](#sec-evalCogMotivational)" or "[intangibles](#sec-evalCogMotivational)"), the scout can score this across multiple players and include these data in the actuarial prediction formula.
For instance, the scout could use a rating scale (e.g., 1 = "poor"; 2 = "fair"; 3 = "good"; 4 = "very good"; 5 = "excellent") to code (i.e., translate) their qualitative judgment into a quantifiable rating that can be integrated with other information in the actuarial formula.
That said, the quality of predictions rests on the quality and relevance of the assessment information for the particular prediction decision.
If the assessment data are lousy, it is unlikely that a statistical algorithm (or a human for that matter) will make an accurate prediction:
"Garbage in, garbage out".
A statistical formula cannot rescue inaccurate assessment data.

>>>>

As an example of using the actuarial approach, Billy Beane, who was the general manager of the Oakland Athletics at the time, wanted to find ways for his team—which had less financial resources than its competitors—to compete with teams that hard more money to sign players.
Because the team did not have the resources to sign the best players, they had to find to find other ways to find the optimal players that they could afford.
So, he used statistical formulas that weight variables, such as on-base percentage and slugging percentage, according to their value for winning games, for the use of selecting players.
His approach became well-known based on the Michael Lewis book, "Moneyball: The Art of Winning an Unfair Game", and the eventual movie, "Moneyball".

@@

There is sometimes a misconception that formulas cannot account for qualitative information.
However, that is not true.
Qualitative information can be scored or coded to be quantified so that it can be included in statistical formulas.
For instance, if an expert scout is able to meaningfully assess a player's [cognitive and motivational factors](#sec-evalCogMotivational) (i.e., the "[X factor](#sec-evalCogMotivational)" or "[intangibles](#sec-evalCogMotivational)"), the scout can score this across multiple players and include these data in the actuarial prediction formula.
For instance, the scout could use a rating scale (e.g., 1 = "poor"; 2 = "fair"; 3 = "good"; 4 = "very good"; 5 = "excellent") to code (i.e., translate) their qualitative judgment into a quantifiable rating that can be integrated with other information in the actuarial formula.
That said, the quality of predictions rests on the quality and relevance of the assessment information for the particular prediction decision.
If the assessment data are lousy, it is unlikely that a statistical algorithm (or a human for that matter) will make an accurate prediction:
"Garbage in, garbage out".
A statistical formula cannot rescue inaccurate assessment data.
Likewise, if the input data are based on measurement strategies that are biased against particular individuals or groups, the actuarial formula is not immune from bias.
